Between Cell Variance

### 2020-05-21 Chap3_MLH1var.rmd
notes and preliminary results for Chap3

load(file="~./MLH1repo/data/Results_setup_4.21.20.RData")

@@

#skeletonize
SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)

@@

@@

Below is the table for the number of cells per category by quality score. For these between cell variance will just use those in quality 1-2.

@@

FULL.male.Var.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(species == "M.musculus") ),
                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Musc male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free")+ylim(c(0,55))

Dom.male.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(subsp == "Dom")  ),
                             aes(color= strain, size = Ncells) ) +geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Dom male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)

mol.male.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(subsp == "Mol")  ),
                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Mol male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)

@@

FULL.female.Var.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(species == "M.musculus") ),
                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Musc male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free")+ylim(c(0,55))

#variance plots
musc.female.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(subsp == "Musc")  ),
                             aes(color= strain, size = Ncells))+
  geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2, guide = FALSE )+ ggtitle("Q12 VAR Musc female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)

Dom.female.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(subsp == "Dom")  ),
                             aes(color= strain, size = Ncells) ) +geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Dom female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)

mol.female.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(subsp == "Mol")  ),
                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Mol female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)

FULL.female.Var.plot
```

# Models (var)

@@

# Models (cV)
```{r cVMLH1co.lmer.MM1, echo=FALSE, waring=FALSE, message=FALSE, eval=TRUE}

@@

cVMLH1.MM_interaction <- drop1(varMLH1.M1, test="Chisq")$`Pr(Chi)`[2]
cVMLH1.MM_sex <- anova(varMLH1.M1, varMLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
cVMLH1.MM_subp <- anova(varMLH1.M1, varMLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
cVMLH1.MM_strain<- MM.rand.pval <- exactRLRT(varMLH1.M1)$p.value

cVMLH1.MMlmer_results <- data.frame(sub.sp = cVMLH1.MM_subp, 
sex.results =  varMLH1.MM_sex, 
               interaction.pval = cVMLH1.MM_interaction,  rand.pval = 
cVMLH1.MM_strain)

@@

# Permutations

@@

#Permutation Step Up

Here's the permutation test I was imagining:
1. Separate the sexes.

@@

8. Compare observed variance in variances to null distribution.

#observed values

<!--
random permutations below, (COMPUTE the VAS)

For the permuted data set, **pool all cells within a strain (permuted strain ie defined by sample size?) and compute the variance for each strain**.

@@

# Permutation Output

```{r permutations2, echo=FALSE}
#I made a DF of strain X sex specific variance

@@

#sampling.size <- nrepeat-3
#careful aboout matching the size of the dataframe!!

list.VAS_F <- c()
for(ff in 1:length(VAS.F.permuted.DF[,2:9997]) ){
  list.VAS_F[ff]  <- var( VAS.F.permuted.DF[,ff] )
}
#list of VAS
list.VAS_F.DF <- data.frame( "female.VAS" =  list.VAS_F)

@@

>>>>

#load(file="~./MLH1repo/data/Results_setup_4.21.20.RData")

@@

#skeletonize
#SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)

Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_DF.xlsx", header=TRUE, 1)

@@

#Writing Outline

I find these results encouraging. The higher variance in females is visible across a variety of genetic backgrounds, including those in which females have lower means than males.

I feel like this result solidifies a plan for Chapter 3.

It could take the following basic form:
Lower stringency of checkpoints during oogenesis predicts higher variance in crossover number within females
Despite acceptance of this prediction, it has rarely been tested
We tested the prediction across a wide array of genetic backgrounds
Females consistently have higher variance than males in the same inbred line, with little evidence for line effects
[Perhaps something about variance in SC length]
In terms of recombination mechanisms, our results mean ...
In terms of evolutionary mechanisms, our results imply ...

@@

Below is the table for the number of cells per category by quality score. For this document just the cells with quality 1-2 will be used.

@@

FULL.male.Var.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(species == "M.musculus") ),
                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain), width = .2, alpha=0.5 )+ ggtitle("Q12 VAR male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free")+ylim(c(0,55))

@@

FULL.female.Var.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(species == "M.musculus") ),
                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain), width = .2, alpha=0.5 )+ ggtitle("Q12 VAR female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free", ncol = 2)+ylim(c(0,55))

FULL.female.Var.plot

#variance plots
musc.female.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(subsp == "Musc")  ),
                             aes(color= strain, size = Ncells))+
  geom_jitter(aes(x = as.factor(strain), y =var, color= strain), alpha=0.5, width = .2, guide = FALSE )+ ggtitle("Q12 VAR Musc female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)

# observed values

# Mean and Variance correlations

```{r var.x.mean, echo=FALSE, warning=FALSE, message=FALSE}

#FULL_mouse_table_Q12.MLH1
Fcor.test <- cor.test( (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(species == "M.musculus") )$mean_co,
  (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female")%>% filter(species == "M.musculus") )$var)

Mcor.test <- cor.test( (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(species == "M.musculus")%>% filter(species == "M.musculus"))$mean_co,
  (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(species == "M.musculus"))$var)

small.DF <- data.frame( corr = paste("corr =", c(Fcor.test$estimate, Mcor.test$estimate) ), 
                     p.val = paste("p.val =", c(Fcor.test$p.value, Mcor.test$p.value) ), sex = c("female", "male") )

Cor.plot <- ggplot(( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus") ), aes(x=mean_co, y=var, color = strain)) + geom_point() + facet_wrap(~sex)+scale_color_manual(values=colors_of_strains, guide = FALSE) +  geom_text(data= small.DF, aes(x=27, y=80, label=corr),  colour="black", inherit.aes=FALSE, parse=FALSE ) +  geom_text(data= small.DF, aes(x=27, y=75, label=p.val),  colour="black", inherit.aes=FALSE, parse=FALSE )+ggtitle("Mean by CO variance per Mouse")

Cor.plot
```

# Models MLH1(var)

@@

# Models MLH1(cV)

```{r cVMLH1co.lmer.MM1, echo=FALSE, waring=FALSE, message=FALSE, eval=TRUE}
#these are all mouse level values

@@

cVMLH1.MM_interaction <- drop1(cVMLH1.M1, test="Chisq")$`Pr(Chi)`[2]
cVMLH1.MM_sex <- anova(cVMLH1.M1, cVMLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
cVMLH1.MM_subp <- anova(cVMLH1.M1, cVMLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
cVMLH1.MM_strain<- MM.rand.pval <- exactRLRT(cVMLH1.M1)$p.value

cVMLH1.MMlmer_results <- data.frame(sub.sp = cVMLH1.MM_subp, 
sex.results =  cVMLH1.MM_sex, 
               interaction.pval = cVMLH1.MM_interaction,  rand.pval = 
cVMLH1.MM_strain)

@@

# Permutations; MLH1

@@

<!--
Here's the permutation test I was imagining:
1. Separate the sexes.

@@

@@

```{r permutations, echo=FALSE}
#I made a DF of strain X sex specific variance

@@

#careful aboout matching the size of the dataframe!!
#calculate the variance across permutations
list.VAS_F <- c()
for(ff in 1:length(VAS.F.permuted.DF[,2:9997]) ){
  list.VAS_F[ff]  <- var( VAS.F.permuted.DF[,ff] )
}
#list of VAS
list.VAS_F.DF <- data.frame( "female.VAS" =  list.VAS_F)

@@

# Permutations: var(F) - var(M)

1. Choose 1 strain.
2. Compute femaleVar-maleVar (pooling all high-quality cells within sexes for that strain).
3. Permute cells between sexes.
4. Compute femaleVar-maleVar for the permuted dataset.
5. Repeat 3-4 10,000 times.
6. Compare the observed value to the distribution of permuted values to compute a p-value.
7. Repeat 1-6 separately for each strain.

```{r Obs.SDvar, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, message=FALSE}

#observed sex difference in varince, by strain
#Q12.MLH1_data by strain

Diff.Var  <- ddply(Q12.MLH1_data, c("strain"), summarise,

                  N.F.mice = length( unique( mouse[which(sex == "female")])  ),
                  N.F.cells  = length( unique(fileName[which(sex == "female")]) ),

                  N.M.mice = length(unique(mouse[which(sex == "male")]) ),
                  N.M.cells = length(unique(fileName[which(sex == "male")]) ),

                F.mean.MLH1 = mean( nMLH1.foci[which(sex == "female")]),
                F.var.MLH1 = var(nMLH1.foci[which(sex == "female")]),

                M.mean.MLH1 = mean( nMLH1.foci[which(sex == "male")]),
                M.var.MLH1 = var(nMLH1.foci[which(sex == "male")]),

               Fvar.Mvar.dif = F.var.MLH1 - M.var.MLH1

)

Diff.Var <- add_subsp(Diff.Var)
Diff.Var <- add_species(Diff.Var)

#plot the values of difference

Var.axis.plot <- ggplot( data = (Diff.Var %>% filter(species == "M.musculus") ), aes(x=F.var.MLH1, y=M.var.MLH1, color=strain))+geom_point()+scale_color_manual(values=colors_of_strains) + ylim(c(0,30))+xlim(c(0,30))+ ggtitle("MLH1 variance by Sex")
Var.axis.plot

Dif.Var.plot <- ggplot( data = (Diff.Var %>% filter(species == "M.musculus") ), aes(x=strain, y=Fvar.Mvar.dif, color=strain))+geom_point()+scale_color_manual(values=colors_of_strains, guide = FALSE) + ggtitle("Degree of Sex Difference in MLH1 Variance")

Dif.Var.plot

Above are the plots for observed sex differences for pooled cell variance. The first plot shows that all female values are higher than male, but PWD and KAZ have the lowest female variance and close to the highest male variance.

```{r permute.by.strain, echo=FALSE}

#pool.cells within strain, calc a Fvar and Mvar, calc the difference, plug into a lis / DF

FM.var.diff_permuted <- data.frame( "pwd.FM.dif" = replicate(n=nrepeat, 
              #female var         
             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "PWD")], Diff.Var$N.F.cells[Diff.Var$strain == "PWD"], replace = FALSE) ) -
              #male var
              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "PWD")], Diff.Var$N.M.cells[Diff.Var$strain == "PWD"], replace = FALSE)  )
                        )  ),

             "msm.FM.dif" = replicate(n=nrepeat, 
              #female var         
             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "MSM")], Diff.Var$N.F.cells[Diff.Var$strain == "MSM"], replace = FALSE) ) -
              #male var
              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "MSM")], Diff.Var$N.M.cells[Diff.Var$strain == "MSM"], replace = FALSE)  )
                        ) ),

             "skive.FM.dif" = replicate(n=nrepeat, 
              #female var         
             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "SKIVE")], Diff.Var$N.F.cells[Diff.Var$strain == "SKIVE"], replace = FALSE) ) -
              #male var
              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "SKIVE")], Diff.Var$N.M.cells[Diff.Var$strain == "SKIVE"], replace = FALSE)  )
                        ) ),
             "wsb.FM.dif" = replicate(n=nrepeat, 
              #female var         
             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "WSB")], Diff.Var$N.F.cells[Diff.Var$strain == "WSB"], replace = FALSE) ) -
              #male var
              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "WSB")], Diff.Var$N.M.cells[Diff.Var$strain == "WSB"], replace = FALSE)  )
                        ) ),
             "lew.FM.dif" = replicate(n=nrepeat, 
              #female var         
             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "LEW")], Diff.Var$N.F.cells[Diff.Var$strain == "LEW"], replace = FALSE) ) -
              #male var
              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "LEW")], Diff.Var$N.M.cells[Diff.Var$strain == "LEW"], replace = FALSE)  )
                        ) ),

             "g.FM.dif" = replicate(n=nrepeat, 
              #female var         
             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "G")], Diff.Var$N.F.cells[Diff.Var$strain == "G"], replace = FALSE) ) -
              #male var
              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "G")], Diff.Var$N.M.cells[Diff.Var$strain == "G"], replace = FALSE)  )
                        ) ),

              "kaz.FM.dif" = replicate(n=nrepeat, 
              #female var         
             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "KAZ")], Diff.Var$N.F.cells[Diff.Var$strain == "KAZ"], replace = FALSE) ) -
              #male var
              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "KAZ")], Diff.Var$N.M.cells[Diff.Var$strain == "KAZ"], replace = FALSE)  )
                        ) ),

             "molf.FM.dif" = replicate(n=nrepeat, 
              #female var         
             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "MOLF")], Diff.Var$N.F.cells[Diff.Var$strain == "MOLF"], replace = FALSE) ) -
              #male var
              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "MOLF")], Diff.Var$N.M.cells[Diff.Var$strain == "MOLF"], replace = FALSE)  )
                        ) )

)#dataframe, FM.var.diff_permuted

#transpose the perrmuted DF, add a strain col  
#t_FM.var.diff_permuted <- t(FM.var.diff_permuted)
#Perm2.DF <- data.frame(names = row.names(t_FM.var.diff_permuted), #t_FM.var.diff_permuted); rownames(t_FM.var.diff_permuted) <- NULL
#Perm2.DF <- add_sex_chp3(Perm2.DF) #sex doesn't make sense for this
#Perm2.DF <- add_strain_Chap3(Perm2.DF)

#calc p - values  (permuted > observed )
#FM.var.diff_permuted

```{r plot.perm2, echo=FALSE, fig.width = 4, fig.height=2.5, message=FALSE}

#plot the first permuted DF-- don't need the transposed
pwd.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = pwd.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "PWD"] ) + ggtitle("PWD difference in variance across sexes")  + annotate("text", x=3, y=800, label= "p=" ) + annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$pwd.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "PWD"] ) / nrepeat )  
pwd.perm2.plot

skive.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = skive.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "SKIVE"] ) + ggtitle("SKIVE difference in variance across sexes") + annotate("text", x=3, y=800, label= "p=" ) + annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$skive.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "SKIVE"] ) / nrepeat )
skive.perm2.plot

msm.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = msm.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "MSM"] ) + ggtitle("MSM difference in variance across sexes")+ annotate("text", x=3, y=800, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$msm.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "MSM"] ) / nrepeat )
msm.perm2.plot

g.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = g.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "G"] ) + ggtitle("G difference in variance across sexes")+ annotate("text", x=3, y=800, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$g.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "G"] ) / nrepeat )
g.perm2.plot

wsb.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = wsb.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "WSB"] ) + ggtitle("WSB difference in variance across sexes")+ annotate("text", x=3, y=800, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$wsb.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "WSB"] ) / nrepeat )
wsb.perm2.plot

lew.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = lew.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "LEW"] ) + ggtitle("LEW difference in variance across sexes")+ annotate("text", x=3, y=800, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$lew.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "LEW"] ) / nrepeat )
lew.perm2.plot

kaz.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = kaz.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "KAZ"] ) + ggtitle("KAZ difference in variance across sexes")+ annotate("text", x=3, y=750, label= "p=" ) +annotate("text", x=5, y=750, label= sum(FM.var.diff_permuted$kaz.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "KAZ"] ) / nrepeat )
kaz.perm2.plot

molf.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = molf.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "MOLF"] ) + ggtitle("MOLF difference in variance across sexes")+ annotate("text", x=3, y=750, label= "p=" ) +annotate("text", x=9, y=750, label= sum(FM.var.diff_permuted$molf.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "MOLF"] ) / nrepeat )
molf.perm2.plot

The above plots show the difference of Female.var - Male.var for n=10,000 permutations.

For all strains the female cell variance is larger than if cells are randomized across sex. But the degree of the difference varies.  PWD and KAZ females have smaller between cell variance.

I should assess how these strain differences might be affected by the number of mice and cells, (espically MOLF).

# Total SC

I isolated cells with Q12 with MLH1 count and Skel measure. I removed rows that had a 'yes' for redo crop.

```{r mouse.total.SC, echo=FALSE}

# factor -> chr
Total.SC_DF$n <- as.character(Total.SC_DF$n)
Total.SC_DF$nMLH1.foci<- as.character(Total.SC_DF$nMLH1.foci)
Total.SC_DF$quality<- as.character(Total.SC_DF$quality)

Total.SC_DF$bin_size<- as.character(Total.SC_DF$bin_size)
Total.SC_DF$skel_size<- as.character(Total.SC_DF$skel_size)

#chr -> num
Total.SC_DF$n <- as.numeric(Total.SC_DF$n)
Total.SC_DF$nMLH1.foci<- as.numeric(Total.SC_DF$nMLH1.foci)
Total.SC_DF$quality<- as.numeric(Total.SC_DF$quality)

Total.SC_DF$bin_size<- as.numeric(Total.SC_DF$bin_size)
Total.SC_DF$skel_size<- as.numeric(Total.SC_DF$skel_size)

Q12.mouse.level.skel <- ddply(.data= (Total.SC_DF %>% filter(quality < 3) ),
                          .(mouse),
                          summarize, 
                          ncells = length(unique(Original.Name)),
                          nmice = length(unique(mouse)),
                          mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
                          var.MLH1 = var(nMLH1.foci),

                          mean.skel = mean(skel_size),
                          var.skel = var(skel_size),

                          mean.bin = mean(bin_size)
)

Q12.mouse.level.skel <- add_strain(Q12.mouse.level.skel)
Q12.mouse.level.skel <- add_sex(Q12.mouse.level.skel)
Q12.mouse.level.skel <- add_category(Q12.mouse.level.skel)
Q12.mouse.level.skel <- add_subsp(Q12.mouse.level.skel)
Q12.mouse.level.skel <- add_species(Q12.mouse.level.skel)

mouse.level.skel <- ddply(.data= (Total.SC_DF %>% filter(quality < 5) ),
                          .(mouse),
                          summarize, 
                          ncells = length(unique(Original.Name)),
                          nmice = length(unique(mouse)),

                          Q12MLH1.mean = mean( nMLH1.foci[which(quality < 3)], na.rm = TRUE ),
                          Q12MLH1.var = var( nMLH1.foci[which(quality < 3)], na.rm = TRUE ),
            #  mean(chromosomeLength[which(hand.foci.count == 1)]
                          Q12skel.mean = mean( skel_size[which(quality < 3)], na.rm = TRUE ),
                          Q12skel.var = var( skel_size[which(quality < 3)], na.rm = TRUE ),

                          mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
                          var.MLH1 = var(nMLH1.foci),

                          mean.skel = mean(skel_size),
                          var.skel = var(skel_size),

                          mean.bin = mean(bin_size)
)

mouse.level.skel <- add_strain(mouse.level.skel)
mouse.level.skel <- add_sex(mouse.level.skel)
mouse.level.skel <- add_category(mouse.level.skel)
mouse.level.skel <- add_subsp(mouse.level.skel)
mouse.level.skel <- add_species(mouse.level.skel)

```{r matching, echo=FALSE}

#plot the differences between the larger dataset and Q12

ii <- ggplot(mouse.level.skel, aes(x=Q12MLH1.mean, y=mean.MLH1, color=strain, size = ncells))+geom_point()+scale_color_manual(values=colors_of_strains, guide = FALSE)

skel.skel <- ggplot(mouse.level.skel, aes(x=Q12skel.mean, y=mean.skel, color=strain, size = ncells))+geom_point()+scale_color_manual(values=colors_of_strains, guide = FALSE)

skel.skel.var.F <- ggplot( data = (mouse.level.skel %>% filter(species == "M.musculus") %>% filter(sex == "female") ), aes(x=Q12skel.var, y=var.skel, color=strain, size = ncells))+geom_point()+scale_color_manual(values=colors_of_strains, guide = FALSE)+facet_wrap(~ subsp, scales = "free", ncol=2)+ylim(c(0, 310000))+xlim(c(0, 310000))
#for some reason, Q12 has a greater range -- but might just be a outlier

#woah, sex differences are large. there is one female PWD with really large variance, 17apr15_PWD_f4 (11 cells)

skel.skel.var.all <- ggplot( data = (mouse.level.skel %>% filter(species == "M.musculus") ), aes(x=Q12skel.var, y=var.skel, color=strain, size = ncells))+geom_point()+scale_color_manual(values=colors_of_strains, guide = FALSE)+xlim(c(0, 500000))+ylim(c(0, 500000))

```{r tot.sc.plot, echo=FALSE, warning=FALSE, message=FALSE}
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
#mouse.level.skel  #mouse level mlh1 and skel
#table(mouse.level.skel$sex)

#
total.SC.plots.F <- ggplot( ( mouse.level.skel %>% filter(sex == "female")%>% filter(species == "M.musculus") %>% filter(strain != "CAST" )  ),  aes(color= strain, size = ncells))+geom_jitter(aes(x = as.factor(strain), y = var.skel, color= strain, alpha=0.5), width = .2 )+ggtitle("Variance of Total Skel Female by Mouse")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free", ncol=2)+ ylim(c(0, 350000 ) )

total.SC.plots.M <- ggplot( ( mouse.level.skel %>% filter(sex == "male")%>% filter(species == "M.musculus") ),  aes(color= strain, size = ncells))+geom_jitter(aes(x = as.factor(strain), y = var.skel, color= strain, alpha=0.5), width = .2 )+ggtitle("Variance of Total Skel Male by Mouse")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free")+ ylim(c(0, 180000 ) )

total.SC.plots.F
total.SC.plots.M

This is the number of cells per category

`r kable(table((Total.SC_DF %>% filter(quality < 3) )$sex, (Total.SC_DF %>% filter(quality < 3) )$strain ))`

These plots are for the number of mice `r kable(table(Q12.mouse.level.skel$sex, Q12.mouse.level.skel$strain))`

```{r corr.total.skel, echo=FALSE, warning=FALSE, message=FALSE}

#FULL_mouse_table_Q12.MLH1
tot.SC.Fcor.test <- cor.test( (Q12.mouse.level.skel %>% filter(sex == "female") %>% filter(species == "M.musculus") )$mean.skel,
  (Q12.mouse.level.skel %>% filter(sex == "female")%>% filter(species == "M.musculus") )$var.skel)

tot.SC.Mcor.test <- cor.test( (Q12.mouse.level.skel %>% filter(sex == "male") %>% filter(species == "M.musculus")%>% filter(species == "M.musculus"))$mean.skel,
  (Q12.mouse.level.skel %>% filter(sex == "male") %>% filter(species == "M.musculus"))$var.skel)

total.SC.small.DF <- data.frame( corr = paste("corr =", c(tot.SC.Fcor.test$estimate, tot.SC.Mcor.test$estimate) ), 
                     p.val = paste("p.val =", c(tot.SC.Fcor.test$p.value, tot.SC.Mcor.test$p.value) ), sex = c("female", "male") )

total.SC.Cor.plot <- ggplot(( Q12.mouse.level.skel %>% filter(species == "M.musculus") ), aes(x=mean.skel, y=var.skel, color = strain)) + geom_point() + facet_wrap(~sex)+scale_color_manual(values=colors_of_strains, guide = FALSE) +  geom_text(data= total.SC.small.DF, aes(x=2000, y=300000, label=corr),  colour="black", inherit.aes=FALSE, parse=FALSE ) +  geom_text(data= total.SC.small.DF, aes(x=2000, y=250000, label=p.val),  colour="black", inherit.aes=FALSE, parse=FALSE )+ggtitle("Mean by total SC variance per Mouse")

# Total SC Models of mouse-level variance

Changing the quality cut off I apply for the total SC dataset changes the significance of the sex coefficient.

```{r total.SC.model.set.up, echo=FALSE, eval=T, warning=FALSE}

#Total.SC_DF - cells
#sex-specific   Q12.mouse.level.skel

Q12.mouse.level.skel$subsp <- factor(Q12.mouse.level.skel$subsp, ordered = FALSE)

Q12.mouse.level.skel$sex <- factor(Q12.mouse.level.skel$sex, ordered = FALSE)

Q12.mouse.level.skel$strain <- factor(Q12.mouse.level.skel$strain, ordered = FALSE)

mouse.level.skel$subsp <- factor(mouse.level.skel$subsp, ordered = FALSE)
mouse.level.skel$sex <- factor(mouse.level.skel$sex, ordered = FALSE)
mouse.level.skel$strain <- factor(mouse.level.skel$strain, ordered = FALSE)

```{r total.SC.var.lmer.MM1, echo=FALSE, waring=FALSE, message=FALSE, eval=T}

#Q12.mouse.level.skel
varTot.SC.M1 <- lmer(var.skel ~ subsp * sex + (1|strain), data= ( mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")  ) )

varTot.SC.M1.reduce.sex <- lmer( var.skel ~ subsp  +(1|strain), data= 
                                   ( mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")  ) 
                                 )

varTot.SC.M1.reduce.subp <- lmer( var.skel ~ sex  +(1|strain), data= ( mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC") ) )

#summary(varTot.SC.M1)$coefficients

varTot.SC.MM_interaction <- drop1(varTot.SC.M1, test="Chisq")$`Pr(Chi)`[2]

varTot.SC.MM_sex <- anova(varTot.SC.M1, varTot.SC.M1.reduce.sex)$`Pr(>Chisq)`[2]

varTot.SC.MM_subp <- anova(varTot.SC.M1, varTot.SC.M1.reduce.subp)$`Pr(>Chisq)`[2]

varTot.SC.MM_strain <- exactRLRT(varTot.SC.M1)$p.value

varTot.SC.MMlmer_results <- data.frame(  sub.sp = varTot.SC.MM_subp, 
        sex.results =  varTot.SC.MM_sex, 
               interaction.pval = varTot.SC.MM_interaction,  rand.pval = 
varTot.SC.MM_strain)

**M1** Mixed model result for the mouse level for: subspecies, sex, interaction and strain. The p values for the total sc variance are as follow, `r varTot.SC.MMlmer_results`. There are significant subspecies, sex and interaction effects, (but no significant strain effects). This uses the larger dataset for total SC.

```{r total.SC.varMLH1.glm.M2, echo=FALSE, eval=T}

#strain = fixed, 
totalSC.M2.glm.strain.fixed.full <- glm(var.skel ~ subsp * sex * strain, data= ( Q12.mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")   ) )

summ.totalSC.M2.glm.strain.fixed.full <- summary(totalSC.M2.glm.strain.fixed.full)
